---
title: "Accessing the OncoKB API with oncoKBData"
authors: "Marcel Ramos Perez"
date: "`r format(Sys.time(), '%B %d, %Y')`"
vignette: >
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteIndexEntry{oncoKBData User Guide}
  %\VignetteEncoding{UTF-8}
output:
  BiocStyle::html_document:
    number_sections: no
    toc: yes
    toc_depth: 4
---

```{r, setup, include=FALSE}
knitr::opts_chunk$set(cache = TRUE)
```

# oncoKBData

The aim of the package is to expose the OncoKB API through an R client.
This vignette demonstrates public API access. To learn more about the OncoKB
database, visit https://www.oncokb.org.

# Installation

To get the development version of `oncoKBData` use: 

```{r,eval=FALSE}
if (!requireNamespace("BiocManager", quietly = TRUE))
    install.packages("BiocManager")

BiocManager::install("waldronlab/oncoKBData")
```

# Package Load

```{r,include=TRUE,results="hide",message=FALSE,warning=FALSE}
library(oncoKBData)
library(S4Vectors)
```

# Introduction

The `oncoKBData` aims to provide access to the OncoKB API via the public
API. Access is also possible with a licensed token.

## API representation

In order to use the OncoKB API, we must instantiate an API object as provided
by the `r CRANpkg("rapiclient")` and `r Biocpkg("AnVIL")` packages.

```{r}
oncokb <- oncoKB()
```

Note that for private API access, users must change the `api.` argument
in the `oncoKB` function.

## Operations

Check available tags, operations, and descriptions as a `tibble`:

```{r}
tags(oncokb)
head(tags(oncokb)$operation)
```

__Note__. The annotations API access requires a token. 

## Levels of Evidence

To retrieve the levels of evidence for all types (i.e., 'therapeutic',
'diagnostic', 'prognostic', and 'FDA') run the `levelsOfEvidence` function. 

```{r}
(loe <- levelsOfEvidence(oncokb))
```

It will return a `DataFrame` with important `metadata`:

```{r}
names(metadata(loe))

metadata(loe)["oncoTreeVersion"]

metadata(loe)[["apiVersion"]]
```

## Gene tables

The API allows retrieval of curated genes where there is a single gene per
observation: 

```{r}
curatedGenes(oncokb)
```

and a long list of genes associated with cancer where there can be multiple
entries for the same `hugoSymbol` due to multiple `geneAliases`:

```{r}
cancerGeneList(oncokb)
```

# Session Information

```{r}
sessionInfo()
```

